---
title: "a2_task2_kat_leigh"
author: "Kat Leigh"
date: "1/31/2021"
output:
  html_document:
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(ggtext)
library(GGally)
library(broom)
library(jtools)

```
use binary logistic regression to test feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens or Sabal etonia. Use code folding and hide all messages & warnings in your knitted HTML. 


Data source: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

More information and metadata: https://portal.edirepository.org/nis/metadataviewer?packageid=edi.317.1
```{r}
# read in data


palmetto <- read_csv("palmetto.csv") %>% clean_names() %>% 
  select(height, length, width, green_lvs, species, plant) %>% 
  mutate(species = ifelse(species == 1, "Serenoa repens",
                          ifelse(species == 2, "Sabal etonia", 0)))

# look at data
look <- palmetto %>%
  select(species, height:green_lvs) %>% 
  ggpairs(aes(color = species))

look
```

## binary logistic regression

```{r}
palmetto_blr <- glm(as.factor(species) ~ height + length + width, green_lvs, 
                            data = palmetto, 
                            family = "binomial")

# Look at the model: 

palmetto_blr

summary(palmetto_blr)

# Get a tidy version w/ broom: 
blr_tidy <- broom::tidy(palmetto_blr)

# above are the coefficients for log odds

# convert to probabilities

blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")

```
Vizualize findings

```{r}

ggplot(data = blr_fitted, aes(x = green_lvs, y = .fitted)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species), se = FALSE) +
  labs(x = "Number of green leaves",
       y = "Probability of outcome Sabal etonia")+
  theme_minimal()

# Visualization of p(Sabal etonia) by variable

# For length:
effect_plot(palmetto_blr, 
            pred = length, 
            interval = TRUE, 
            y.label = "Probability of 'Sabal etonia'")

# For height:
effect_plot(palmetto_blr, 
            pred = height, 
            interval = TRUE, 
            y.label = "Probability of 'Sabal etonia'")

# for width
effect_plot(palmetto_blr, 
            pred = width, 
            interval = TRUE, 
            y.label = "Probability of 'Sabal etonia'")

# for green leaves
effect_plot(palmetto_blr, 
            pred = green_lvs, 
            interval = TRUE, 
            y.label = "Probability of 'Sabal etonia'")


# Predictions for new values with `predict()`

What is the probability that a plant of  height XXXX with a length of XXXX, width of XXXX, and XXXX number of green leaves will be Sabal etonia?

ex_1 <- predict(palmetto_blr, 
                data.frame(height = 120, 
                length = 250, 
                width = 200),
                green_lvs = 15)

# Based on the model, the probability that this plant is a Sabal etonia is XXXX.
```

You can also feed in a new data frame, with multiple penguin observations, to get model probability estimates for more than one plant: 

```{r}
new_df <- data.frame(
  height = c(70, 120, 170),
  length = c(110, 170, 220),
  width = c(100, 160, 210),
  green_lvs = c(7, 12, 17)
)

ex_2 <- predict(palmetto_blr,
                new_df,
                type = "response")

```

